---
title: "Tidy Tuesday 2018 Week 01"
author: "Thomas J. Demmer"
date: "1/20/2019"
output: html_document
---

The following analysis of average tuition 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message=FALSE}
# Standard workspace setup
remove(list = ls())
options(stringsAsFactors = FALSE)
options(scipen = 999)
```

Libraries:
```{r message=FALSE}
library(readxl)
library(tidyverse)
library(ggthemes)
library(gganimate)
library(gifski)
library(viridis)
library(gridExtra)

```

Import data:
```{r warning=FALSE}
df <- read_excel("~/rstats/rviz/tidy_tuesday_2018_week_01/us_avg_tuition.xlsx")
```

View dim of data
```{r}
dim(df)
```

View top five rows of dataframe
```{r}
head(df)
```

Grab the region from our state_region excel file in our code tab and use inner_join to add the region to our code, then select our data and view the head to make sure it appears how we want it.
```{r}
df_region <- df %>% 
  inner_join(read_excel("~/rstats/code/us_states_regions.xlsx") %>% 
               select(Region, State)) %>% 
  select(State, Region, `2004-05`:`2015-16`)
head(df_region)
```

Use the tidyr function gather to change dataframe from wide to long form and arrange by state.  View top five rows of dataframe in long format.
```{r}
df_long <- df_region %>%
  gather(Year, Tuition, -State, -Region)
head(df_long)
```

Prepare first graph using geom_jitter
```{r}
df_long %>% 
  ggplot(aes(x = Year, y = Tuition, color = Region)) +
  geom_jitter(width = 0.11, size = 5, alpha = 0.7, shape = 19) +
  theme_minimal() +
  scale_color_brewer(type = "div", palette = 5) +
  #scale_color_viridis(option = "D", discrete = TRUE, begin = 0.6) +
  theme(legend.position = "bottom")
```

```{r}
library(ggridges)
```

```{r}
df_long %>% 
  ggplot(aes(x = Tuition, y = fct_rev(Year), fill = Region)) +
  geom_density_ridges_gradient(color = "black",
                               show.legend = FALSE,
                               alpha = 0.3,
                               panel_scaling = TRUE,
                               quantile_lines = TRUE, quantiles = 2) +
  facet_wrap(~fct_reorder(Region, Tuition, mean), scales = "free") +
  scale_x_continuous(limits = c(1000, 18000),
                     breaks = seq(0, 180000, 3000),
                     labels = scales::comma) +
  theme_minimal() +
  labs(x = "Average tuition",
       y = "Academic year",
       title = "Distribution of average tuition by region",
       subtitle = "Solid line denotes the median",
       caption = "Tidy tuesday week 1: Tuition price by state") +
  scale_fill_viridis(alpha = 0.9, option = "C", discrete = TRUE, begin = 0.3) +
  theme(panel.grid.minor.x = element_blank(),
        plot.title = element_text(size = 15, face = "bold"),
        plot.subtitle = element_text(size = 12),
        strip.text = element_text(size = 10, face = "bold"))
```

#Trying gg highlights
```{r}
track_by_region <- df_long %>% 
  group_by(Region, Year) %>% 
  summarise(Tuition = median(Tuition)) %>% 
  mutate(year_code = as.numeric(factor(Year)))

df_long %>% 
  mutate(year_code = as.numeric(factor(Year))) %>% 
  ggplot(aes(x = year_code, y = Tuition, group = State)) +
  geom_line(show.legend = FALSE, color = "gray", size = 3, alpha = 0.5) +
  geom_line(aes(x = year_code, y = Tuition, color = Region),
            data = track_by_region, inherit.aes = FALSE, size = 5, alpha = 0.5) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.grid.minor.x = element_blank(),
        axis.title.x = element_text(size = 13, face = "bold"),
        axis.text.x = element_text(size = 11),
        axis.title.y = element_text(size = 13, face = "bold"),
        axis.text.y = element_text(size = 11),
        legend.title = element_text(face = "bold"),
        plot.title = element_text(face = "bold", size = 15),
        plot.subtitle = element_text(size = 13)) +
  scale_x_continuous(limits = c(1, 12),
                     breaks = seq(1, 12, by = 1),
                     labels = unique(df_long$Year)) +
  scale_y_continuous(limits = c(2000, 16000),
                     breaks = seq(2000, 16000, by = 2000),
                     labels = scales::comma) +
  labs(x = "\nYear",
       y = "Median tuition price",
       title = "Change in median tuition price",
       subtitle = "Median regional changes highlighted") +
  guides(color = guide_legend(
    title.position = "top",
    title.hjust = 0.5,
    label.position = "top",
    keywidth = 4
  )) +
  viridis::scale_color_viridis(alpha = 0.9, option = "C", discrete = TRUE)
```



Mutate Year show first four numbers (remove the "-XX" from the year)
```{r}
#df_long2 <- df_long %>%
#  mutate(Year = str_sub(Year, start = 1, end = 4))

#head(df_long2)
```



Graph states with lines (geom_smooth) over time
```{r}
ggplot(df_long2, aes(x = Year,
                          y = Tuition,
                          color = State)) +
  geom_smooth(aes(x = Year, y = Tuition, group = State),
              se = FALSE,
              method = "loess",
              formula = "y ~ x") +
  scale_y_continuous(labels = scales::comma) +
    theme(legend.position = "none",
        rect = element_rect(fill = "#0a1926"),#0a1926
        panel.grid = element_blank(),
        text = element_text(colour = "white"),
        panel.background = element_rect(fill = "transparent", colour = "transparent"),
        axis.text = element_text(colour = rgb(230, 230, 230, maxColorValue = 255), size = 11)) +
  labs(x = "",
       y = "TUITION",
       title = "STATE TUITION") 
```

Filtering data to show only the stats that I have lived in (NY, NH, MA, VA)
```{r}
df_filtered <- df_long2 %>% 
  filter(State %in% c("Massachusetts", "New York", "New Hampshire", "Virginia"))

df_filtered
```

Graph the 4 states we filtered from above
```{r}
ggplot(df_filtered, aes(x = Year,
                          y = Tuition,
                          color = State)) +
  geom_smooth(aes(x = Year, y = Tuition, group = State),
              se = FALSE,
              method = "loess",
              formula = "y ~ x") +
  geom_text(data = subset(df_filtered, Year == 2015), aes(label = State), vjust = 0.2, hjust = 2) +
  scale_y_continuous(labels = scales::comma) +
  theme(legend.position = "none",
        rect = element_rect(fill = "#0a1926"),#0a1926
        panel.grid = element_blank(),
        text = element_text(colour = "white"),
        panel.background = element_rect(fill = "transparent", colour = "transparent"),
        axis.text = element_text(colour = rgb(230, 230, 230, maxColorValue = 255), size = 11)) +
  labs(x = "",
       y = "TUITION",
       title = "STATE TUITION")
```




